---
title: "How to handle adjusted/delayed columns?"
author: "Dan Kelley"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

**Abstract.** This notebook, made while thinking about `useBest()`, examines
the data for a particular float, downloaded from
<https://data-argo.ifremer.fr/dac/aoml/5904013/profiles/D5904013_153.nc>, which
had 3 columns, with `DATA_MODE` of DAD, meaning that columns 1 and 3 are
delayed, while column 2 is adjusted.

**Question.** What a user might try to do with those 3 columns?  For example,
are all of them useful, or just the 'adjusted' one?

# Load the data

We read `DATA_MODE` with netcdf, and then (as `dataMode`, now a vector) with
`read.oce()`.

```{r}
library(ncdf4)
library(oce)
library(curl)

url <- "https://data-argo.ifremer.fr/dac/aoml/5904013/profiles/D5904013_153.nc"
file <- gsub(".*/", "", url)
if (!file.exists(file))
    curl_download(url, file)

n <- nc_open(file)
ncvar_get(n, "DATA_MODE")
d <- read.oce(file)
d[["dataMode"]]
```

So, we are reading `dataMode` correctly.  (This is not a surprise, but it was
worth demonstrating.)

## Define flag-interpretation functions

Note that these functions handle both single variables (say for S only) or
pairs of variables (say S and T, or S, T and p).

In oce, flags are described as follows (based on descriptions from an official Argo document):
`not_assessed`=0, `passed_all_tests`=1, `probably_good`=2, `probably_bad`=3, `bad`=4,
`changed`=5, `not_used_6`=6, `not_used_7`=7, `estimated`=8, and `missing`=9.

```{r}
# accept if flagged with 1=passed, 2=probably good, 5=changed
flagIsAcceptable <- function(f1, f2=NULL, f3=NULL)
{
    if (is.null(f2))
        f2 <- rep(2, length(f1))
    if (is.null(f3))
        f3 <- rep(2, length(f1))
    (f1 %in% c(1, 2, 5)) & (f2 %in% c(1, 2, 5)) & (f3 %in% c(1, 2, 5))
}

# black if acceptable, red otherwise
flagCol <- function(f1, f2=NULL, f3=NULL)
{
    if (is.null(f2))
        f2 <- rep(2, length(f1))
    if (is.null(f3))
        f3 <- rep(2, length(f1))
    ifelse(flagIsAcceptable(f1, f2, f3), "black", "red")
}
```

\newpage

## Initial data examination

```{r}
summary(d)
```

The QC tests are as follows
```{r}
d[["historyAction"]] 
d[["historyQCTest"]] 
argoFloats::showQCTests(d)
```

\newpage

## More in-depth data examination

```{r}
p <- d[["pressure"]]
pf <- d[["pressureFlag"]]
padj <- d[["pressureAdjusted"]]
padjf <- d[["pressureAdjustedFlag"]]

S <- d[["salinity"]]
Sf <- d[["salinityFlag"]]
Sadj <- d[["salinityAdjusted"]]
Sadjf <- d[["salinityAdjustedFlag"]]

T <- d[["temperature"]]
Tf <- d[["temperatureFlag"]]
Tadj <- d[["temperatureAdjusted"]]
Tadjf <- d[["temperatureAdjustedFlag"]]

npcol <- ncol(p)
if (npcol != ncol(S))
    stop("p and S have differing numbers of columns")
if (npcol != ncol(T))
    stop("p and T have differing numbers of columns")
pDF <- data.frame(p1=p[,1], p2=p[,2], p3=p[,3],
    pf1=pf[,1], pf2=pf[,2], pf3=pf[,3],
    padj1=padj[,1], padj2=padj[,2], padj3=padj[,3],
    padjf1=padjf[,1], padjf2=padjf[,2], padjf3=padjf[,3])
SDF <- data.frame(S1=S[,1], S2=S[,2], S3=S[,3],
    Sf1=Sf[,1], Sf2=Sf[,2], Sf3=Sf[,3],
    Sadj1=Sadj[,1], Sadj2=Sadj[,2], Sadj3=Sadj[,3],
    Sadjf1=Sadjf[,1], Sadjf2=Sadjf[,2], Sadjf3=Sadjf[,3])
TDF <- data.frame(T1=T[,1], T2=T[,2], T3=T[,3],
    Tf1=Tf[,1], Tf2=Tf[,2], Tf3=Tf[,3],
    Tadj1=Tadj[,1], Tadj2=Tadj[,2], Tadj3=Tadj[,3],
    Tadjf1=Tadjf[,1], Tadjf2=Tadjf[,2], Tadjf3=Tadjf[,3])
head(pDF)
tail(pDF)
head(SDF)
tail(SDF)
head(TDF)
tail(TDF)
```

\newpage

## Plot unadjusted data, red if flagged as unacceptable

```{r fig.height=5.5}
par(mfrow=c(4, npcol), mar=c(3,3,1,1), mgp=c(2,0.7,0))
ctd1 <- as.ctd(S[,1], T[,1], p[,1])
ctd2 <- as.ctd(S[,2], T[,2], p[,2])
ctd3 <- as.ctd(S[,3], T[,3], p[,3])
plotProfile(ctd1, xtype="S", col=flagCol(Sf[,1], pf[,1]), type="p", pch=20)
plotProfile(ctd2, xtype="S", col=flagCol(Sf[,2], pf[,2]), type="p", pch=20)
plotProfile(ctd3, xtype="S", col=flagCol(Sf[,3], pf[,3]), type="p", pch=20)
plotProfile(ctd1, xtype="T", col=flagCol(Tf[,1], pf[,1]), type="p", pch=20)
plotProfile(ctd2, xtype="T", col=flagCol(Tf[,2], pf[,2]), type="p", pch=20)
plotProfile(ctd3, xtype="T", col=flagCol(Tf[,3], pf[,3]), type="p", pch=20)
plotProfile(ctd1, xtype="sigmaTheta", col=flagCol(Sf[,1], Tf[,1], pf[,1]), type="p", pch=20)
plotProfile(ctd2, xtype="sigmaTheta", col=flagCol(Sf[,2], Tf[,2], pf[,2]), type="p", pch=20)
plotProfile(ctd3, xtype="sigmaTheta", col=flagCol(Sf[,3], Tf[,3], pf[,3]), type="p", pch=20)
plotTS(ctd1, col=flagCol(Sf[,1], Tf[,1], pf[,1]))
plotTS(ctd2, col=flagCol(Sf[,2], Tf[,2], pf[,2]))
plotTS(ctd3, col=flagCol(Sf[,3], Tf[,3], pf[,3]))
```

\newpage

## Plot unadjusted data, skipping unacceptable (to narrow scales)

```{r fig.height=5.5}
par(mfrow=c(4, npcol), mar=c(3,3,1,1), mgp=c(2,0.7,0))
accept <- flagIsAcceptable(Sf[,1], Tf[,1], pf[,1])
ctd1 <- as.ctd(S[,1][accept], T[,1][accept], p[,1][accept])
accept <- flagIsAcceptable(Sf[,2], Tf[,2], pf[,2])
ctd2 <- as.ctd(S[,2][accept], T[,2][accept], p[,2][accept])
accept <- flagIsAcceptable(Sf[,3], Tf[,3], pf[,3])
ctd3 <- as.ctd(S[,3][accept], T[,3][accept], p[,3][accept])
plotProfile(ctd1, xtype="S", type="p", pch=20)
plotProfile(ctd2, xtype="S", type="p", pch=20)
plotProfile(ctd3, xtype="S", type="p", pch=20)
plotProfile(ctd1, xtype="T", type="p", pch=20)
plotProfile(ctd2, xtype="T", type="p", pch=20)
plotProfile(ctd3, xtype="T", type="p", pch=20)
plotProfile(ctd1, xtype="sigmaTheta", type="p", pch=20)
plotProfile(ctd2, xtype="sigmaTheta", type="p", pch=20)
plotProfile(ctd3, xtype="sigmaTheta", type="p", pch=20)
plotTS(ctd1)
plotTS(ctd2)
plotTS(ctd3)
```

\newpage

## Plot adjusted data, red if flagged as unacceptable

```{r fig.height=5.5}
par(mfrow=c(4, npcol), mar=c(3,3,1,1), mgp=c(2,0.7,0))
ctd1 <- as.ctd(Sadj[,1], Tadj[,1], padj[,1])
ctd2 <- as.ctd(Sadj[,2], Tadj[,2], padj[,2])
ctd3 <- as.ctd(Sadj[,3], Tadj[,3], padj[,3])
plotProfile(ctd1, xtype="S", col=flagCol(Sadjf[,1], padjf[,1]), type="p", pch=20)
plotProfile(ctd2, xtype="S", col=flagCol(Sadjf[,2], padjf[,2]), type="p", pch=20)
plotProfile(ctd3, xtype="S", col=flagCol(Sadjf[,3], padjf[,3]), type="p", pch=20)
plotProfile(ctd1, xtype="T", col=flagCol(Tadjf[,1], padjf[,1]), type="p", pch=20)
plotProfile(ctd2, xtype="T", col=flagCol(Tadjf[,2], padjf[,2]), type="p", pch=20)
plotProfile(ctd3, xtype="T", col=flagCol(Tadjf[,3], padjf[,3]), type="p", pch=20)
plotProfile(ctd1, xtype="sigmaTheta", col=flagCol(Sadjf[,1], Tadjf[,1], padjf[,1]), type="p", pch=20)
plotProfile(ctd2, xtype="sigmaTheta", col=flagCol(Sadjf[,2], Tadjf[,2], padjf[,2]), type="p", pch=20)
plotProfile(ctd3, xtype="sigmaTheta", col=flagCol(Sadjf[,3], Tadjf[,3], padjf[,3]), type="p", pch=20)
plotTS(ctd1, col=flagCol(Sadjf[,1], Tadjf[,1], padjf[,1]))
plotTS(ctd2, col=flagCol(Sadjf[,2], Tadjf[,2], padjf[,2]))
plotTS(ctd3, col=flagCol(Sadjf[,3], Tadjf[,3], padjf[,3]))
```

\newpage

## Plot adjusted data, skipping if flagged as unacceptable (to improve scales)

```{r fig.height=5.5}
par(mfrow=c(4, npcol), mar=c(3,3,1,1), mgp=c(2,0.7,0))
accept <- flagIsAcceptable(Sadjf[,1], Tadjf[,1], padjf[,1])
ctd1 <- as.ctd(Sadj[,1][accept], Tadj[,1][accept], padj[,1][accept])
accept <- flagIsAcceptable(Sadjf[,2], Tadjf[,2], padjf[,2])
ctd2 <- as.ctd(Sadj[,2][accept], Tadj[,2][accept], padj[,2][accept])
accept <- flagIsAcceptable(Sadjf[,3], Tadjf[,3], padjf[,3])
ctd3 <- as.ctd(Sadj[,3][accept], Tadj[,3][accept], padj[,3][accept])
plotProfile(ctd1, xtype="S", type="p", pch=20)
plotProfile(ctd2, xtype="S", type="p", pch=20)
plotProfile(ctd3, xtype="S", type="p", pch=20)
plotProfile(ctd1, xtype="T", type="p", pch=20)
plotProfile(ctd2, xtype="T", type="p", pch=20)
plotProfile(ctd3, xtype="T", type="p", pch=20)
plotProfile(ctd1, xtype="sigmaTheta", type="p", pch=20)
plotProfile(ctd2, xtype="sigmaTheta", type="p", pch=20)
plotProfile(ctd3, xtype="sigmaTheta", type="p", pch=20)
plotTS(ctd1)
plotTS(ctd2)
plotTS(ctd3)
```

