---
title: "Introduction to argoFloats Quality Control"
author: "Jaimie Harbin (https://orcid.org/0000-0003-3774-3732), Dan Kelley (https://orcid.org/0000-0001-7808-5911),and Clark Richards (https://orcid.org/0000-0002-7833-206X)"
date: "`r Sys.Date()`"
---

**Abstract**

Argo floats undergo a series of testing to ensure the data provided is a
reliable source. Understanding the Quality Control (QC) completed on profiles
can be problematic, however. The creators of `argoFloats` therefore developed a
way to deal with flags and adjusted data. When dealing with flags, a three step
process exists involving `plot(which='QC')`, `showQCTests()`and `applyQC()`. In
addition to flag control, `argoFloats` also deals with adjusted data with the
`useAdjusted()` function.

# Introduction to Flags

Core, biogeochemical (BGC), and Deep Argo data all undergo testing to ensure the
data found at the Data Assembly Centers are as accurate as possible. More
specifically, testing is done in three levels: 1: Real time that complete
checks on all measurements and assigned a quality flag within a 24-48 hour
time frame, 2. Delayed mode, and 3. Regional scientific analyses of all float
data with other available data. The procedures for regional analyses are still
to be determined.

At first glance, it can be difficult for users to understand the Quality Control
(QC) completed on profiles. To address this issue when dealing with Argo data,
`argoFloats` created a way to deal with flags and adjusted data. Firstly, the
creators of  `argoFloats` designed a three step process to deal with flags. Step
one is `plot(which='QC')`, which shows the quality and mean of parameters. Step
two is `showQCTests()`, which demonstrates which QC tests were performed and
failed. Step three is `applyQC()`, which allows the users to get rid of 'bad'
data if they wish. In addition to flag control, `argoFloats` also deals with
adjusted data in `useAdjusted()`, which allows users to use adjusted data if
available.

# Flags

As described by the Argo user’s manual [2], data that are given a flag value of
1, 2, 5, or 8 are considered 'good' data. If data are given a flag value of 9
(or missing value) they are not used, and flags 3, 5, 6, and 7 are considered
'bad' data.

`argoFloats` has taken this idea, and has incorporated a 3 step process when
considering the quality control of Argo flags. The three step process encourages
users to always analyze the data they are considering. The three step process is
as follows:

```{r qualityControlProcess, echo=FALSE, fig.cap="Figure 1: Work flow of the argoFloats package.", fig.height=1.8, fig.width=5, fig.align='center', dev.args=list(pointsize=10), warning=FALSE}
colText <- "darkblue"
colCode <- "black"
library(graphics)
textInBox <- function(x, y, text, cex=1, pos=4, center=TRUE, family="Times New Roman", col="black", tweakx=0)
{
    w <- graphics::strwidth(text)
    h <- graphics::strheight(text)
    if (center)
        x <- x - w/2
    text(x+tweakx, y, text, cex=cex, pos=pos, family=family, col=col)
    rect(x+tweakx, y-h, x+tweakx+1.1*w, y+h, border=col)
    invisible(list(w=w, h=h))
}
omar <- par("mar")
par(mar=c(0,1,0,0))
plot(c(0, 1), c(0, 1), type="n", xlab="", ylab="", axes=FALSE)
x0 <- 0.25
y0 <- 0.9
dy <- 0.3
wh <- textInBox(x0, y0, "    What is the quality of the data?", col=colText)
h <- wh$h * 1.2
xarrow <- x0 + 0.01 # not sure why we need to move it
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "   What tests were performed?", col=colText)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "   Set low-quality data to NA.", col=colText)
x0 <- 0.8
y0 <- 0.9
wh <- textInBox(x0, y0, "plot(which='QC') ", family="sans", col=colCode)
h <- wh$h * 1.2
xarrow <- x0 + 0.01 # not sure why we need to move it
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, "showQCTests() ", family="sans", col=colCode)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, "applyQC()  ", family="sans", col=colCode)
par(mar=omar)
```

This vignette will walk the users through the three step process using the float with
ID 1901584 near the Bahamas.

### Step One: QC Plot

Our `argoFloats` package uses the flags of profiles to determine "good" and
"bad" data. We then created a plot type to demonstrate the quality of data, the
`QC` plot. The QC plot is a plot of parameter quality and parameter mean. This
only works if `x` is an object that was created by `getProfiles()`. The user
must also provide the `parameter` of interest. If the user is uncertain about
what parameter exists within a certain float, `argoFloats` will prompt the user
with the appropriate parameters.

An example of how to get the QC plot for temperature, or "TEMP" for the float with ID 1901584 is as follows:

```{r, warning=FALSE, error=FALSE, message=FALSE}
library(argoFloats)
data("index")
subset <- subset(index, ID='1901584')
profiles <- getProfiles(subset)
argos <- readProfiles(profiles)
plot(argos, which='QC', parameter='temperature')
```

*Exercise One:* Use the previous code to determine the QC of temperature for
float number 4900845 in synthetic data.Explain your results.

### Step Two: showQCTests()

The `showQCTests()` is a newly added function. It's purpose is to use integer
values from the `hexToNibble()` function to internally convert hex digits. In
this context, the hex digits come from `HISTORY_QCTEST` of a single argoFloats
object that was created by `readProfiles()`. The `showQCTests()` is then used to
indicate which QC tests were performed and/or failed.

Using the same example as shown in `plot(which='QC')`, we can also determine
which QC tests were performed on each profile. This is completed by
`showQCTests()` in the following code:

```{r, message=FALSE, error=FALSE, warning=FALSE}
library(argoFloats)
data('index')
subset <- subset(index, 1)
profiles <- getProfiles(subset)
argos <- readProfiles(profiles)
argos1 <- argos[[1]]
showQCTests(argos[[1]])
```

In section 3.11 Reference table 11: QC test binary IDs of the Argo User's Manual
[2], we see that test two is the "Impossible data test".

*Exercise Two*: Use the first float with ID 4900845 in exercise one,to determine
which tests were performed and/or failed during quality control testing. EXPLAIN
YOUR RESULTS.

### Step Three: applyQC()

The `applyQC()` function examines quality-control within an argoFloats
object that was created by `readProfiles()`. By default, it replaces all suspicious
data with NA values, so they will not appear in plots or be considered in
calculations. This is an important early step in processing, because suspicious
Argo floats commonly report data that are suspicious based on statistical and
physical measures.

Using the same example as shown in `plot(which='QC')` and `showQCTests()`, the
user also has the ability to use `applyQC()` as shown below:

```{r, message=FALSE, error=FALSE, warning=FALSE}
library(argoFloats)
# Contrast TS diagrams for raw and flag-handled data
data(index)
i <- subset(index, ID='1901584')
raw <- readProfiles(getProfiles(i))
clean <- applyQC(raw)
par(mfrow=c(1, 2))
plot(raw, which="TS")
plot(clean, which="TS")
```

*Exercise Three* : Use `applyQC()` on float ID with 4900845. Explain your results.

# Adjusted Data

### useAdjusted()

In addition to flags flagging bad data based on tests, `argoFloats` package has
also incorporated the idea of `<param>Adjusted` vs `<param>Unadjusted` which
considers "bad" data that has been adjusted [2].

Variables with original names indicating in the string `<param>Adjusted` are
assigned nicknames names ending in Adjusted by `readProfiles()`, so that e.g.
`doxyAdjusted` gets the nickname `oxygenAdjusted`, while `doxy` gets the
nickname `oxygen`. `useAdjusted()` switches these, renaming the adjusted values,
so that e.g. `DOXY_ADJUSTED` gets nickname `oxygen` and `DOXY` gets nickname
`oxygenUnadjusted`. This is carried out for all data families, and also for the
corresponding units and quality-control flags. See `?useAdjusted` to understand
when `argoFloats` chooses to use adjusted data.

An example of the significance of adjusted vs unadjusted can be shown below
using the ID from the previous flag examples:

```{r, warning=FALSE, error=FALSE, message=FALSE}
library(argoFloats)
data("index")
subset <- subset(index, ID='1901584')
profiles <- getProfiles(subset)
raw <- readProfiles(profiles)
adj <- useAdjusted(raw)
par(mfrow=c(1, 2))
oce::plotProfile(oce::as.ctd(raw[[1]]), xtype="temperature")
mtext("Raw data", side=3, line=-1, col=2)
oce::plotProfile(oce::as.ctd(adj[[1]]), xtype="temperature")
mtext("Adjusted data", side=3, line=-1, col=2)
```

*Exercise Four* Use `useAdjusted()` on the float with ID5903586, profile=50 to
compare `<param>Adjusted` vs `<param>Unadjusted>`.

# Answers

*Exercise One:* Use the previous code to determine the QC of temperature for
float number 4900845 in synthetic data. Explain your results.
```
library(argoFloats)
ais <- getIndex(filename = 'synthetic', age=0)
sub <- subset(ais, ID='4900845')
profiles <- getProfiles(sub)
argos <- readProfiles(profiles)
plot(argos, which='QC', parameter='temperature')
```
As shown in the plot, 100% of the temperature data is considered "good". This
implies to the user that all data is considered okay to use.

*Exercise Two*: Use the first float with ID 4900845 in exercise one,to determine
which tests were performed and/or failed during quality control testing. Explain your results.

```
library(argoFloats)
data('indexSynthetic')
sub <- subset(indexSynthetic, ID='4900845')
profiles <- getProfiles(sub)
argos <- readProfiles(profiles)
a1 <- argos[[1]]
showQCTests(a1)
```
As shown by the `showQCTests()` function, no quality control tests were actually
performed on this float ID. This is a case where the user would have to self
analyze their data to ensure all data seems reasonable.

*Exercise Three* : Use `applyQC()` on float ID = 4900845. Explain your results.
```
library(argoFloats)
# Contrast TS diagrams for raw and flag-handled data
data(indexSynthetic)
i <- subset(index, ID='4900845')
raw <- readProfiles(getProfiles(i))
clean <- applyQC(raw)
par(mfrow=c(1, 2))
plot(raw, which="TS")
plot(clean, which="TS")
```
As shown by the plot, the plot containing "good" data and the plot containing
"bad" data are the same. This is the result of no tests being completed for QC, as shown by `showQCTests()`.

*Exercise Four* Use `useAdjusted()` on the float with ID 5903586, profile=50 to
compare `<PARAM>_ADJUSTED` vs `<param>Unadjusted`.

```{r}
library(argoFloats)
bai <- getIndex('synthetic')
s <- subset(bai, ID='5903586')
ss <- subset(s, 50)
profiles <- getProfiles(ss)
raw <- readProfiles(profiles)
adj <- useAdjusted(raw)
par(mfrow=c(1, 2))
oce::plotProfile(oce::as.ctd(raw[[1]]), xtype="oxygen")
mtext("Raw data", side=3, line=-1, col=2)
oce::plotProfile(oce::as.ctd(adj[[1]]), xtype="oxygen")
mtext("Adjusted data", side=3, line=-1, col=2)
```

# References
1. Wong, A., Keeley, R. and Carval, T., 2009. Argo quality control manual. Version, 2(5), p.32.
2. Carval, Thierry, Bob Keeley, Yasushi Takatsuki, Takashi Yoshida, Stephen Loch Loch, Claudia Schmid, and Roger Goldsmith. Argo User’s Manual V3.3. Ifremer, 2019. https://doi.org/10.13155/29825.
