---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

# argoFloats

<!-- badges: start -->

[![Project Status: Active – The project has reached a stable, usable
state and is being actively
developed.](http://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/argoFloats)](https://cran.r-project.org/package=argoFloats)
[![R build status](https://github.com/ArgoCanada/argoFloats/workflows/R-CMD-check/badge.svg)](https://github.com/ArgoCanada/argoFloats/actions)
![RStudio CRAN mirror downloads](https://cranlogs.r-pkg.org/badges/last-month/argoFloats)
![RStudio CRAN mirror downloads](https://cranlogs.r-pkg.org/badges/last-week/argoFloats)
![RStudio CRAN mirror downloads](https://cranlogs.r-pkg.org/badges/last-day/argoFloats)
![RStudio CRAN mirror downloads](https://cranlogs.r-pkg.org/badges/grand-total/argoFloats)

<!-- badges: end -->

The argoFloats package for the R language provides tools for downloading and
analyzing collections of oceanographic Argo float datasets.  Its developers are
Dan Kelley (Dalhousie University), Jaimie Harbin (Bedford Institute of
Oceanography and Dalhousie University), Clark Richards (Bedford Institute of
Oceanography) and Dewey Dunnington (Bedford Institute of Oceanography).

The CRAN version may be installed with
```{r eval=FALSE}
install.packages("argoFloats")
```

The development version may be installed with
```{r eval=FALSE}
library(devtools)
install_github("ArgoCanada/argoFloats", ref = "develop")
```
where, of course, the `devtools` package must be installed first, if it is not already present on the user's system.

Typical use will require also installing several other packages, using
```{r eval=FALSE}
install.packages(c(
    "shinyBS", "colourpicker", "curl", "lubridate", "ncdf4",
    "oce", "ocedata", "s2", "sf", "shiny"
))
```

Once things are set up as above, it will be possible to use all the features of
`argoFloats`, many of which are illustrated in the documentation for its
functions, and in the vignette that is built into the package, both of which
are displayed on the [user-oriented
website](https://argocanada.github.io/argoFloats/index.html). Note that the
vignette also lists Youtube videos about the package.

As a practical example, the code block given below shows how to create a map
and a temperature-salinity diagram for Argo float profiles near Abaco Island in
the Bahamas. The key steps, many of which are common to most analyzes using the
package, are as follows.

1. The `getIndex` function is used to download a worldwide index of float
   profiles.  (Use `?getIndex` to learn more about this function, and note in
particular the `destdir` argument, which determines where the index and other
argo files will be stored locally.)

2. The `subset` function is used to narrow the region of interest, yielding 39
   profiles as of mid-June, 2020.

3. `getProfiles` is used to download the NetCDF files that contain the profile
   measurements.

4. `readProfiles` is used to read those files.

5. `applyQC` is used to set questionable data to `NA`.  This uses
   quality-control (QC) flags that are present in the argo data.

6. `plot` is used to create a map plot.  The image shows water depth in metres,
   determined through a call to the `getNOAA.bathy` function in the `marmap`
package. For context, a star is drawn to indicate the centre of he focus
region.  A label is drawn above the map, showing the number of profiles.  This
illustrates the use of `[[`, a generic R function that is used in argoFloats to
access data and metadata elements in argoFloats objects, as well as things that
can be computed from these elements, such as seawater density, etc.  Use
`?"[[,argoFloats-method"` to see the details of how `[[` works.

7. `plot` is used again, to make a temperature-salinity diagram.

With this in mind, readers ought to find it easy to read the following code.  A
reasonable next step would be to try altering the code, perhaps to explore
another region or to see whether the QC step is actually necessary.

```{r results="hide", echo=FALSE}
oldpar <- par(no.readonly = TRUE)
```

```{r results="hide", out.width="80%", fig.height=3, fig.width=5, dev.args=list(pointsize=8)}
library(argoFloats)
library(oce)
## 1. Get worldwide float-profile index, saving to ~/data/argo by default.
indexAll <- getIndex()
## 2. Narrow to a 30km-radius circle centred on Abaco Island, The Bahamas.
index <- subset(indexAll,
    circle = list(longitude = -77.06, latitude = 26.54, radius = 30)
)
## 3. Get NetCDF files for these profiles, saving to ~/data/argo by default.
profiles <- getProfiles(index)
## 4. Read the NetCDF files.
argos <- readProfiles(profiles)
## 5. Examine QC flags, and set questionable data to NA.
argosClean <- applyQC(argos)
par(mfrow = c(1, 2)) # want two-panel plot
par(mar = c(3.5, 2.0, 2.0, 2.0)) # maps do not get axis names
par(mgp = c(2, 0.7, 0)) # tighten axes
## 6. Plot a map of profile locations.
plot(index, which = "map", bathymetry = FALSE)
points(-77.06, 26.54, pch = "*", cex = 3) # show centre of focus
mtext(paste(argosClean[["length"]], "profiles"), line = 1.0)
## 7. Plot a TS diagram
par(mar = c(3.5, 3.5, 2.0, 1.0)) # increase left margin for name
plot(argosClean, which = "TS")
```

```{r results="hide", echo=FALSE}
par(oldpar)
```



