---
title: "Tweet Thread"
output: html_document
---

# Tweet One

Hi Argo community! I want to introduce you to a newly developed package R package, argoFloats. This package was created by myself, Dan Kelley, and Jaimie Harbin to provide tools for downloading and analyzing collections of oceanographic Argo float datasets.

**Insert argoFloats logo**


# Tweet Two
argoFloats has an easy-to-follow work flow, to allow the users to effectively access, download, and read Argo data.

```{r workflow, echo=FALSE, fig.cap="Figure 1: Work flow of the argoFloats package.", fig.height=2.8, fig.width=5, fig.align='center', dev.args=list(pointsize=10), warning=FALSE}
colText <- "darkblue"
colCode <- "black"
library(graphics)
textInBox <- function(x, y, text, cex=1, pos=4, center=TRUE, family="Times New Roman", col="black", tweakx=0)
{
    w <- graphics::strwidth(text)
    h <- graphics::strheight(text)
    if (center)
        x <- x - w/2
    text(x+tweakx, y, text, cex=cex, pos=pos, family=family, col=col)
    rect(x+tweakx, y-h, x+tweakx+1.1*w, y+h, border=col)
    invisible(list(w=w, h=h))
}
omar <- par("mar")
par(mar=c(0,1,0,0))
plot(c(0, 1), c(0, 1), type="n", xlab="", ylab="", axes=FALSE)
x0 <- 0.25
y0 <- 0.9
dy <- 0.2
wh <- textInBox(x0, y0, "Get index from server", col=colText)
h <- wh$h * 1.2
xarrow <- x0 + 0.01 # not sure why we need to move it
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "Focus on a subset of profiles", col=colText)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "Get profile data files from server", col=colText)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "Read profile data files", col=colText)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "Process the data", col=colText)
x0 <- 0.8
y0 <- 0.9
wh <- textInBox(x0, y0, "getIndex() ", family="sans", col=colCode)
h <- wh$h * 1.2
xarrow <- x0 + 0.01 # not sure why we need to move it
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, "subset() ", family="sans", col=colCode)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, "getProfiles() ", family="sans", col=colCode)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, "readProfiles() ", family="sans", col=colCode)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, ". . .   ", family="sans", col=colCode, tweakx=0.005)
par(mar=omar)
```

# Tweet Three
To get familiar with argoFloats, check out the user website, https://argocanada.github.io/argoFloats/index.html, or developer website https://github.com/ArgoCanada/argoFloats for more advanced R users.

# Tweet Four
A series of real-time examples exists at our youtube channel https://www.youtube.com/channel/UCmVBNwRRGx5sRa1skvfOrvA, which include some East coast humour to enjoy.

# Tweet Five
The following code produces a TS plot near Bermuda:

library(argoFloats)
ai <- getIndex()
ais <- subset(ai, circle=list(longitude=-64.7505, latitude=32.3078, radius=50))
pr <- getProfiles(ais)
a <- readProfiles(pr, handleFlags = TRUE)
plot(a, which="TS")

# Tweet Six
The following link provides code for subsetting by ocean and polygon near the Isthmus of Panama:
https://github.com/ArgoCanada/argoFloats/tree/develop/videos/04

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(argoFloats)
ai <- getIndex()
s <- subset(ai, circle=list(longitude=-83, latitude=9, radius=500))
plot(s, which='map') # To get a visual
atlantic <- subset(s, ocean='A') # Subsetting for Atlantic Ocean
pacific <- subset(s, ocean='P')
points(atlantic[['longitude']], atlantic[['latitude']], pch=20, col=2)
points(pacific[['longitude']], pacific[['latitude']], pch=20, col=3)
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(argoFloats)
ai <- getIndex()
s <- subset(ai, circle=list(longitude=-83, latitude=9, radius=500))
plot(s, which='map', bathymetry=FALSE) # To get a visual
pax <- c(-83.92029, -84.29865, -84.73106, -83.86623, -82.73114, -81.32580, -79.43398,-77.59622, -78.29889,-79.86640,-82.19063)
pay <- c(13.715453, 12.701096, 10.832545, 9.924962, 9.070767, 8.536896, 9.177542, 10.779157, 13.234968, 13.662066, 13.822227)
# Now select your points on the map in the Atlantic
a <- subset(s, polygon=list(longitude=pax, latitude=pay))
points(a[['longitude']], a[['latitude']], col=2)
ppx <- c(-86.13641, -85.05537, -83.70408, -82.19063, -80.51502, -78.78536, -78.56915, -79.32588,-81.16364, -82.46089, -86.24451, -88.51469, -89.05521, -88.35254, -86.62288)
ppy <- c(11.846901, 10.565609,  9.658026,  8.643670,  8.430121,  7.842862,  6.935280,  5.547213, 4.586244,  4.105759,  4.586244,  6.348021,  9.070767, 11.206255, 12.273999)
# Now select your points on the map in the Pacific
p <- subset(s, polygon=list(longitude=ppx, latitude=ppy))
points(p[['longitude']], p[['latitude']], col=3)
```