---
title: "Tweet Thread"
output: html_document
---

# Tweet One

Hi Argo community! I want to introduce you to a newly developed R package, argoFloats. This package was created by myself, Dan Kelley, and Jaimie Harbin to provide tools for downloading and analyzing collections of oceanographic Argo float datasets.

**Insert argoFloats logo found at https://github.com/ArgoCanada/argoFloats/blob/develop/logo/logo.R**


# Tweet Two
argoFloats has an easy-to-follow work flow, to allow the users to effectively access, download, and read Argo data.

```{r workflow, echo=FALSE, fig.cap="Figure 1: Work flow of the argoFloats package.", fig.height=2.8, fig.width=5, fig.align='center', dev.args=list(pointsize=10), warning=FALSE}
colText <- "darkblue"
colCode <- "black"
library(graphics)
textInBox <- function(x, y, text, cex=1, pos=4, center=TRUE, family="Times New Roman", col="black", tweakx=0)
{
    w <- graphics::strwidth(text)
    h <- graphics::strheight(text)
    if (center)
        x <- x - w/2
    text(x+tweakx, y, text, cex=cex, pos=pos, family=family, col=col)
    rect(x+tweakx, y-h, x+tweakx+1.1*w, y+h, border=col)
    invisible(list(w=w, h=h))
}
omar <- par("mar")
par(mar=c(0,1,0,0))
plot(c(0, 1), c(0, 1), type="n", xlab="", ylab="", axes=FALSE)
x0 <- 0.25
y0 <- 0.9
dy <- 0.2
wh <- textInBox(x0, y0, "Get index from server", col=colText)
h <- wh$h * 1.2
xarrow <- x0 + 0.01 # not sure why we need to move it
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "Focus on a subset of profiles", col=colText)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "Get profile data files from server", col=colText)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "Read profile data files", col=colText)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "Process the data", col=colText)
x0 <- 0.8
y0 <- 0.9
wh <- textInBox(x0, y0, "getIndex() ", family="sans", col=colCode)
h <- wh$h * 1.2
xarrow <- x0 + 0.01 # not sure why we need to move it
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, "subset() ", family="sans", col=colCode)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, "getProfiles() ", family="sans", col=colCode)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, "readProfiles() ", family="sans", col=colCode)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, ". . .   ", family="sans", col=colCode, tweakx=0.005)
par(mar=omar)
```

# Tweet Three
To get familiar with argoFloats, check out the user website, https://argocanada.github.io/argoFloats/index.html, or developer website https://github.com/ArgoCanada/argoFloats for more advanced R users.

# Tweet Four
Users can easily sift through data based on geographical region, parameter, time, institution, deep Argo, ID, ocean, mode, and profile. A series of real-time examples exists at our Youtube channel https://www.youtube.com/channel/UCmVBNwRRGx5sRa1skvfOrvA.

# Tweet Five
For example, the following code demonstrates how to use the easy-to-follow work flor to produces a TS plot near Bermuda:

```{r, message=FALSE, warning=FALSE}
library(argoFloats)
# Get world index
data(index)
# Get index withing 50km of Bermuda
ais <- subset(index, 1:10)
# Get profile data files for this near-Bermuda index
pr <- getProfiles(ais)
# Read those data files
a <- readProfiles(pr, handleFlags=TRUE)
# Plot a map to the left, and a TS diagram to the right
par(mfrow=c(1, 2))
plot(a, which="map")
plot(a, which='TS')
```

# Tweet Six
The following link provides code for subsetting by ocean near the Isthmus of Panama. Note it is a subset by ocean "Area", which is why we incorporated the subset by polygon function.
https://github.com/ArgoCanada/argoFloats/tree/develop/videos/04

```{r, warning=FALSE, error=FALSE, message=FALSE}
library(argoFloats)
ai <- getIndex()
s <- subset(ai, circle=list(longitude=-83, latitude=9, radius=500))
plot(s, which='map') # To get a visual
atlantic <- subset(s, ocean='A') # Subsetting for Atlantic Ocean
pacific <- subset(s, ocean='P')
points(atlantic[['longitude']], atlantic[['latitude']], pch=20, col=2)
points(pacific[['longitude']], pacific[['latitude']], pch=20, col=3)
```

# Tweet 07
Code found at https://github.com/ArgoCanada/argoFloats/blob/develop/sandbox/jlh/08_polygonPlot/polygonPlot.R demonstrates how to use our subset by polygon and TS plot function to create a T-S diagram comparing the Atlantic and Pacific Ocean near the Isthmus of Panama.

`**INSERT PHOTO from https://github.com/ArgoCanada/argoFloats/blob/develop/sandbox/jlh/08_polygonPlot/polygonPlot.R**

# Tweet 07
The code found at https://github.com/ArgoCanada/argoFloats/blob/develop/videos/05/video05_TS_by_oxygen.R demonstrates how to use the argoFloats package to create a TS diagram, colour-coded by oxygen.

```{r, message=FALSE, warning=FALSE, error=FALSE, echo=FALSE}
library(oce)
library(argoFloats)

# Acquire merged argo index, subset by circle, and then subset by oxygen
ai <- getIndex(filename='merge')
subc <- subset(ai, circle=list(longitude=-77.5, latitude=27.5, radius=300))
subo <- subset(subc, parameter='DOXY')

# Download argo files, read them, and thn extract O2
profiles <- getProfiles(subo)
argos <- readProfiles(profiles)
O2 <- unlist(argos[["oxygen"]]) # note that [[ returns a list

# Plot colour palette, and then the TS plot
cm <- colormap(O2)
drawPalette(colormap=cm, zlab="Oxygen", pos=3)
plot(argos, which="TS", col=cm$zcol)
```

# Tweet 08
Bathymetry has recently been added to the argoFloats package. If you visit https://github.com/ArgoCanada/argoFloats/blob/develop/videos/07/video07_map_with_bathymetry.R you can find code to create a trajectory plot, colour coded by time, with bathymetry.

```{r, message=FALSE, warning=FALSE, error=FALSE, echo=FALSE}
library(argoFloats)
library(oce)
ID0 <- '6900388'
if (!exists("index0"))
    index0 <- subset(getIndex(), ID=ID0)

lon <- index0[["longitude"]]
lat <- index0[["latitude"]]
profile <- index0[['profile']]
t <- index0[["date"]]

par(mar=c(2,2,1,1))

cm <- colormap(t, col=oceColorsJet)
drawPalette(colormap=cm, tformat="%Y-%m", pos=3)
plot(index0, bathymetry=TRUE)
points(lon, lat, pch=21, cex=1, bg=cm$zcol, col="black", lwd=1.2)
lines(lon, lat)
sub <- seq(1, length(lon), by=10)
text(lon[sub], lat[sub], profile[sub], cex=2/3, pos=1)
mtext(paste("Float", ID0), adj=0)
```