---
title: "Introduction to argoFloats"
author: "Dan Kelley (https://orcid.org/0000-0001-7808-5911), Jaimie Harbin ( https://orcid.org/0000-0003-3774-3732) and Clark Richards (https://orcid.org/0000-0002-7833-206X)"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 5
    fig_height: 5
    dpi: 72
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Introduction to argoFloats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding[utf-8]{inputenc}
---

<!-- HOW TO BUILD AND VIEW THE VIGNETTES.
1. Edit this file in vignettes/argoFloats.Rmd
2. Do not bother trying to build vignettes in RStudio. Instead,
   in a unix shell, type the following:
       R CMD build --compact-vignettes="qpdf" argoFloats
       R CMD install argoFloats_0.1.1.tar.gz
   where the version number may need adjustment in the second command.
3. To see the results, type the following in an R console:
      browseVignettes("argoFloats")
-->


```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```
**Abstract.** The `argoFloats` package makes it easy to get, read, and analyze core, biogeochemical(bgc), and deep argo data. Our goal in making this package was to eliminate the gap between the freely available argo data and the end user.

# Introduction

The `argoFloats` package provides tools for downloading and processing Argo profile data. It allows users to focus on core, biogeochemical (bgc) , or deep argo profiles, and also allows the user to sift these profiles based on ID, time, geography, and institution. Once downloaded, such datasets can be analysed within `argoFloats` or using other R tools and packages.

# Preliminary setup
To set up a computer to use `argoFloats`, the user should first install the following packages (*Hint:* you may have to use the `Sys.setenv(R_REMOTES_NO_ERRORS_FROM_WARNINGS = TRUE)` command if you receive a Zero Exit Status Error):
```{r eval=FALSE}
devtools::install_github("dankelley/oce", ref='develop')
devtools::install_github('dankelley/ocedata', ref='develop')
devtools::install_github('dankelley/argoFloats', ref='develop')
```
Note that the `install_github()` function is provided by the `devtools` package, so the above will fail if that package has not already been installed.  To install it, use
```{r eval=FALSE}
install.packages(devtools)
```

# Work flow

The figure below illustrates the typical workflow with the package, with descriptions of the steps on the left, and names of the relevant functions on the right.
```{r workflow, echo=FALSE, fig.cap="Work flow of the argoFloats package.", fig.height=2.5, fig.width=7, fig.align='center', dev.args=list(pointsize=10)}
colText <- "darkblue"
colCode <- "black"
library(graphics)
textInBox <- function(x, y, text, cex=1, pos=4, center=TRUE, family="Times New Roman", col="black")
{
    w <- graphics::strwidth(text)
    h <- graphics::strheight(text)
    if (center)
        x <- x - w/2
    text(x, y, text, cex=cex, pos=pos, family=family, col=col)
    rect(x, y-h, x+1.1*w, y+h, border=col)
    invisible(list(w=w, h=h))
}
omar <- par("mar")
par(mar=c(0,1,0,0))
plot(c(0, 1), c(0, 1), type="n", xlab="", ylab="", axes=FALSE)
x0 <- 0.25
y0 <- 0.9
dy <- 0.25
wh <- textInBox(x0, y0, "Get index from server", col=colText)
h <- wh$h * 1.2
xarrow <- x0 + 0.01 # not sure why we need to move it
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "Get profile data files from server", col=colText)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "Read profile data files", col=colText)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1, col=colText)
y0 <- y0 - dy
textInBox(x0, y0, "Process the data", col=colText)
x0 <- 0.8
y0 <- 0.9
dy <- 0.25
wh <- textInBox(x0, y0, "getIndex() ", family="sans", col=colCode)
h <- wh$h * 1.2
xarrow <- x0 + 0.01 # not sure why we need to move it
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, "getProfiles() ", family="sans", col=colCode)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, "readProfiles() ", family="sans", col=colCode)
arrows(xarrow, y0-h, xarrow, y0-dy+h, length=0.1)
y0 <- y0 - dy
textInBox(x0, y0, " ...    ", family="sans", col=colCode)
par(mar=omar)
```

As illustrated, the central functions are `getIndex()`, `subset()`, `subset,argoFloats-method()`, `getProfiles()`, and `readProfiles()`. 

Some built-in datasets are also provided for concreteness of illustration and for testing, but actual work always starts with a call to `getIndex()` to download a full index of float data, which we will demonstrate later in this vignette.

# argoFloats Functions

To begin to get familiar with how the `argoFloats` package works, we will begin looking at the built in data sets. Built into the `argoFloats` package is the `index`, `indexBgc`, and `indexMerged`, referring to core argo, bgc-argo, and a combinaton respectively. For the sake of this vignette we will focus on the `indexMerged` data set. 

The first step is to access the required packages that will be needed during this tutorial, with
```{r message=FALSE}
library(oce)
library(ocedata)
library(argoFloats)
```

To access the embedded data set within `argoFloats`, the following code is used:

```{r}
data('indexMerged')
```

It's now possible to apply the  `argoFloats`  specialized versions of R "generic" functions, `plot()`, `[[`, `summary()`, and `show()`.

## Plotting the Data
The `plot()` function within the `argoFloats` package provides simple ways to plot aspects of `argoFloats-class` objects. To produce the built in plot and visualize the coordinates of a section of argo floats off of the Bahamas, the following code is used:

```{r}
plot(indexMerged)
```

## Extracting Data 
`[[` provides a way to extract items from `argoFloats` objects, without getting lost in the details of storage. (Note that `[[<-` is *not* specialized, since the user is highly discouraged from altering values within `argoFloats` objects.). For example, if the user wanted to extract the `parameters` within the `indexMerged` dataset, instead of doing `indexMerged@data$index$parameters`, instead they can simply do `indexMerged[['parameters']]`

## Summarizing Data

The `summary()` command displays key features of `argoFloats-class` objects. See `summary,argoFloats-method()` further details.
 
## Showing Information About argoFloats Objects
Lastly, `show()` command provides a one-line sketch of `argoFloats-class` objects. For example if the user types in:

```
indexMerged

```
The following output occurs:

```
argoFloats object of type "index" with 39 items
```

See `show,argoFloats-method()` for details.

It should be noted that the profile elements within `argoFloats` objects are stored as in the form of `argo` objects as defined by the `oce` package. This means that `argoFloats` users can rely on a wide variety of `oce` functions to analyse their data. The full suite of R tools is also available, and the vastness of that suite explains why `argoFloats` is written in R.
 
 
# Steps to working with data that is not built in

##  getIndex()
As shown by Figure 1, the `getIndex()` command is used to get an index of available argo float profiles, either by downloading information from a data repository or by reusing an index (stored as an .rda file) that was prepared by a recent call to the function.

The `getIndex()` command works by specifying the server, with first trying the USGODAE (ftp://usgodae.org/pub/outgoing/argo) and then the IRREMER (ftp://ftp.ifremer.fr/ifremer/argo) if that doesn't work. The next step is to specify the file name. As shown in the table below, the user has the ability to write the specific file name, which are listed when using the `?getIndex()` function, or to simply use the embedded nicknames within the package, 'argo', 'bgc' or 'bgcargo', and 'merge' or 'merged'. To further help the user, and avoid them needing to specifically locate the server, a description of the contents found within each file name are displayed as well.

File Name                         | Nickname                | Contents
----------------------------------|-------------------------|--------------------------------------
`ar_greylist.txt`                 | -                       | Suspious/malfunctioning floats
`ar_index_global_meta.txt.gz`     | -                       | Metadata files
`ar_index_global_prof.txt.gz`     | `"argo"`                | Argo data
`ar_index_global_tech.txt.gz`     | -                       | Technical files
`ar_index_global_traj.txt.gz`     | -                       | Trajectory files
`argo_bio-profile_index.txt.gz`   | `"bgc"` or `"bgcargo"`  | Biogeochemical data (without S or T)
`argo_bio-traj_index.txt.gz`      | -                       | Biogochemica trajectory files
`argo_merge-profile_index.txt.gz` | `"merge"` or `"merged"` | Merged `"argo"` and `"bgc"` data


See `?getIndex()` for further description about the further arguments.

To get the indexMerged index from the USGODAE server, the following code is used:

```{r}
ai <- getIndex(file = "merge", destdir = "~/data/argo")
```

## subset()

As shown by Figure 1, the next step when working with the `argoFloats` package is to use the `subset()` function to focus on a subset of profiles. The `argoFloats` package provides tools to sift through profiles based on ID, time, geography and instituion. 

If, for example, the user wanted to plot the position of floats within a 100 km radius of the Bermuda, the following code is used:

```{r, message=FALSE}
# Subsetting by circle
Belon <- -64.7505 # Longitude of Bermuda
Belat <- 32.3078 # Latitude of Bermuda
aiBermuda <- subset(ai, circle=list(longitude=Belon, latitude=Belat, radius=100))
# Plotting circle subset
latbe <- aiBermuda[['latitude']]
lonbe <- aiBermuda[['longitude']]
belat <- 32
belon <- -64.5
data(coastlineWorldFine, package="ocedata")
plot(coastlineWorldFine, clongitude=belon, clatitude=belat, col="pink", span=400)
points(lonbe,latbe)
```

*Exercise 1:* Produce the same plot as shown above, however, only include data from 2010.

In addition to the `circle` subset function, the argoFloats package also has the ability to subset by `rectangle` and `polygon`. The `rectangle` subset can be useful when analyzing the section data found within the oce package, which includes a westward transect from the Mediterranean outflow region across to North America. To view this data the following code is used:

```{r fig.cap="Section plot made by oce package.", fig.height=4, fig.width=7, fig.align='center', dev.args=list(pointsize=10)}
data(section, package='oce')
plot(section)
```

*Exercise 2:* Create a rectangle subset for this transect and plot the positions for the year 2017. 

## getProfiles()

The next step within the `argoFloats` package is the `getProfiles()` function. This takes an index constructed with `getIndex()`, possibly after focusing with `subset,argoFloats-method()`, and creates a list of files to download from the server named in the index. Then these files are downloaded to the `destdir` directory, using filenames inferred from the source filenames. The value returned by `getProfiles()` is suitable for use by `readProfiles()` function.

## readProfiles()

The `readProfiles()` command works with either a list of local netCDF files, or a `argoFloats` object type `profiles`, as created by `getProfiles()`.

The command can be useful for analyzing individual profiles, for example:

```{r}
sub <- subset(ai, 1:2) # To subset for profiles
profiles <- getProfiles(sub, destdir='~/data/argo')
argos <- readProfiles(profiles, handleFlags = TRUE)
pressure <- argos[['argos']][[1]][['pressure']]
temperature <- argos[['argos']][[1]][['temperature']]
plot(temperature, pressure, ylim=rev(range(pressure, na.rm=TRUE)), xlab='Temperature (C)', ylab='Depth (dbar)')
```

*Exercise 3:* Using the profile in the previous example, plot a TS diagram.

# Solutions

*Exercise 1:* Produce the same plot as shown above, however, only include data from 2018.

```{r, eval=FALSE}
#Installing necessary packages
library(devtools)
install_github('dankelley/argoFloats', ref='develop')
install_github("dankelley/oce", ref='develop')
install_github('dankelley/ocedata', ref='develop')
#Subsetting index by circle
ai <- getIndex(file='merge', destdir='~/data/argo')
Belon <- -64.7505 # Longitude of Bermuda
Belat <- 32.3078 # Latitude of Bermuda
aiBermuda <- subset(ai, circle=list(longitude=Belon, latitude=Belat, radius=100))
#Subsetting circle by time
from <- as.POSIXct("2010-01-01", tz="UTC")
to <- as.POSIXct("2018-10-31", tz="UTC")
ait <- subset(aiBermuda, time=list(from=from, to=to))
#Plotting new subset on a map
latbe <- ait[['latitude']]
lonbe <- ait[['longitude']]
belat <- 32
belon <- -64.5
data(coastlineWorldFine, package="ocedata")
plot(coastlineWorldFine, clongitude=belon, clatitude=belat, col="pink", span=400)
points(lonbe,latbe)
```
*Exercise 2:* Create a rectangle subset for this transect and plot the positions for the year 2017.

```{r, eval=FALSE}
library(argoFloats)
library(oce)
data(section, package='oce')
lat0 <- median(section[['latitude']])
lon0 <- median(section[['longitude']])
# Subset by rectangle
ai <- getIndex(file='merge', destdir='~/data/argo')
latRect <- lat0 + c(-2,2)
lonRect <- lon0 + c(-30,30)
air <- subset(ai, rectangle=list(longitude=lonRect, latitude=latRect))
# Subset the rectangle by time
from <- as.POSIXct("2017-01-01", tz="UTC")
to <- as.POSIXct("2017-12-31", tz="UTC")
aiTime <- subset(air, time=list(from=from, to=to))
#Plot this subset
latbe <- aiTime[['latitude']]
lonbe <- aiTime[['longitude']]
belat <- 37
belon <- -35
data(coastlineWorldFine, package="ocedata")
plot(coastlineWorldFine, clongitude=belon, clatitude=belat, col="pink", span=6000)
points(lonbe,latbe)
```

*Exercise 3:* Using the profile in the previous example, plot a TS diagram.

```{r, eval=FALSE}
ai <- getIndex(file='merge', destdir = '~/data/argo')
sub <- subset(ai, 1:2) # To subset for profiles
profiles <- getProfiles(sub, destdir='~/data/argo')
argos <- readProfiles(profiles, handleFlags = TRUE)
pressure <- argos[['argos']][[1]][['pressure']]
temperature <- argos[['argos']][[1]][['temperature']]
salinity <- argos[['argos']][[1]][['salinity']]
ctd<-as.ctd(salinity,temperature,pressure)
plotTS(ctd, eos="unesco")
```
