---
title: Working with previously downloaded files
author: "Dan Kelley"
date: "2023-03-03"
---

Source: oce/sandbox/dek/17_get_local_file/17_get_local_file.Rmd

**Abstract.** This document, aimed at co-developers and not users, demonstrates the new ability of `getIndex()` to deal with previously-downloaded files. This relates to issues [584](https://github.com/ArgoCanada/argoFloats/issues/584) and [585](https://github.com/ArgoCanada/argoFloats/issues/585), and to zoom conversations between me, JH and CR on March 1, 2 and 3 of 2023. One of the changes made in this version is to expand the information provided when `debug>0`, including timing information.  Note in particular that the fastest way to work with a previously-downloaded index is to use the `rda` file that was created by `getIndex()` in a previous call.  It makes sense to use the OS to rename such a file for safe-keeping, because otherwise a new call to `getIndex()` will over-write it.

# Setup

```{r}
library(argoFloats)
```

# Save a .gz file for later tests

```{r}
i <- getIndex(debug=3, keep=TRUE)
i1 <- subset(i, 1)
summary(i1)
```

Let's also find the name of the file that will later get downloaded.

```{r}
dataFile <- paste0("~/data/argo/", gsub(".*/", "", i1[["file"]]))
```

# Can we read a previously downloaded .gz file?

Test typical sequence of actions.

```{r}
GZf <- list.files("~/data/argo", "*.gz", full.names=TRUE)[1]
GZi <- getIndex(GZf, server=NULL, debug=3)
summary(GZi)
GZi1 <- subset(GZi, 1)
summary(GZi1)
dataFile <- paste0("~/data/argo/", gsub(".*/", "", GZi1[["file"]]))
unlink(dataFile)
GZp1 <- getProfiles(GZi1)
Sys.time() - file.info(dataFile)$ctime # should be short (just a few Kb download)
summary(GZp1)
GZa1 <- readProfiles(GZp1)
summary(GZa1)
```

# Can we read a previously downloaded .rda file?

Test typical sequence of actions.

```{r}
RDAf <- list.files("~/data/argo", "*.rda", full.names=TRUE)[1]
RDAi <- getIndex(RDAf, server=NULL, debug=3)
summary(RDAi)
RDAi1 <- subset(RDAi, 1)
summary(RDAi1)
dataFile <- paste0("~/data/argo/", gsub(".*/", "", RDAi1[["file"]]))
unlink(dataFile)
RDAp1 <- getProfiles(RDAi1)
Sys.time() - file.info(dataFile)$ctime # should be short (just a few Kb download)
summary(RDAp1)
RDAa1 <- readProfiles(RDAp1)
summary(RDAa1)
```

